Predicting Internet Path Dynamics and Performance with Machine Learning

### 2017-12-20 paper-tex/paper.tex
Minor corrections and abstract fine-tuning.

	\begin{abstract}
		We study the problem of predicting Internet path dynamics and performance with \texttt{traceroute} measurement and machine learning models. In this work, we first point out the drawbacks of the data processing in one previous work \cite{}, and reprocess the data in several more rational ways. We then conduct the experiments on the reprocessed data with random forest model. With 
		the data analysis and experiment results' insights, we further leverage Long Short-Term Memory (LSTM) model, one kind of recurrent neural network (RNN), and get promising results. We make our reprocessed data and experiment code publicly available\footnote{{https://github.com/felixwzh/CN-group-project}}.
	\end{abstract}

@@

	\section{Introduction}
	Internet paths change frequently due to inter/intra-domain routing changes, load balancing, and even misconfigurations and failures\cite{1}.
	%  V. Paxson. 1996. End-to-end Routing Behavior in the Internet. In Proceedings of the ACM SIGCOMM 1996 Conference (SIGCOMM ’96). ACM, New York, NY, USA.
	Some of these changes can seriously disrupt performance, causing longer round-trip times, congestion, or even loss of connectivity\cite{2}.
	%U. Javed, I. Cunha, D. Cho"nes, E. Katz-Bassett, T. Anderson, and A. Krishnamurthy. 2013. PoiRoot: Investigating the Root Cause of Interdomain Path Changes. In Proceedings of the ACM SIGCOMM 2013 Conference on SIGCOMM (SIGCOMM ’13). ACM, New York, NY, USA, 183–194.
	Thus, it is of great significant to predict Internet path dynamics and performance.

	In the original paper\cite{oripaper}, the authors focus on the problem of predicting Internet path changes and path performance using \texttt{traceroute} measurements. They use the recent route information of paths (route age for paths, route changes in the past, average RTT, etc) to do some predictions of paths. Their are three predict targets: $(i)$ the remaining life time of a route (i.e., the time before a route changes), $(ii)$ the number of route changes in a future timeslot, and $(iii)$ the average RTT  that the path $P$ will experience in the next \texttt{traceroute} measurement.

@@

	The dataset the authors provide with is a full week of Paris- \texttt{traceroute} measurements performed through the M-Lab open
	Internet measurement initiative\footnote{https://www.measurementlab.net/}. They observe more than 450,000 different paths sampled through Paris-\texttt{traceroute} measurements from more than 180 geo-distributed servers. However, most of the paths are not periodically sampled during this week. And only 2,346 paths have at least 100 \texttt{traceroute} measurements during the analyzed week. So the dataset contains corresponding infomation of these 2,346 paths.

	We find some drawbacks in their data process.(see section \ref{sec:dataAna}) And in our own work, firstly we reprocess the provided data in more resonable ways and use the same method (random forest model) mentioned by the original paper\cite{oripaper} to compare the results. Secondly, another machine learning method, LSTM model, is applied to get better performance.

	\section{Related Works}
	In the original paper\cite{oripaper}, the problem they solve is to use random forest model to predict three targets for one path: $(i)$ the remaining life time of a route for one path at time $t$ (i.e., the time before a route changes), namely $R_r(t)$, $(ii)$ the number of route changes for one path $P$ in a future timeslot $T$, namely $rcP_T(t)$  and $(iii)$ the average RTT that the path $P$ will experience in the next \texttt{traceroute} measurement, namely $avgRTT_P(t)$.

@@

	And the last group of 44 features ,referred to as $F_C$ decribes statistical properties of path latency, which is relavent to the prediction of the next \texttt{traceroute} measurement. They calculate the staistical properties of four RTT metrics (average, minimum, maximum and standard deviation) reported from each \texttt{traceroute} measurement.

@@

	\section{Data Analysis}
	\label{sec:dataAna}
	\hui{I think this is an important part, who will take this part?}
	\heng{Plots of temporally differentiated prediction targets are given here to show their behavior patterns in temporal domain, thus supporting the use of LSTMs.}
	With careful review of the data used in the original paper, here we: $(i)$ show the authors had used the data incorrectly in the context of making predictions, and $(ii)$ reveal the temporal properties that the original method failed to employ but perfectly fits LSTM model (long short term memory neural networks), which we will later use to improve the performance.
	\subsection{Justification of Data Processing}
	The original paper \cite{oripaper} made a major mistake on data employment: they implicitly used data from the future to predict events in the future, which is not possible. More specifically, all of their tests on the performance of \textit{NetPerfTrace} depend on statistics calculated out of data of all time, including the data they were to predict. Features like percentiles of route life, for example, take the route life of all routes during the whole week into account, including route $r_i$. However, these features still take part in predicting the residual life of $r_i$. It resembles guessing something you already knew.

@@

	No more solid evidence could be provided to prove the periodicity of these patterns because pattern itself is hard to define. However, these results are already enough to make us hopeful about the performance of LSTMs, which are expert in capturing temporal dependencies, no-matter long-term ones or short-term ones.

@@

	\subsection{Long Short-Term Memory}
	In order to leverage the prediction targets' character of temporal locality, we further use Long Short-Term Memory (LSTM) to do time series modeling. More formally, we take $t-1$ most recent \texttt{traceroute} measurements features $(\mathbf{x_1},\mathbf{x_2},...,\mathbf{x_{t-2}},\mathbf{x_{t-1}})$and the current \texttt{traceroute} measurement feature $\mathbf{x_t}$ together as the input,i.e.,$\mathbf{X}=(\mathbf{x_1},\mathbf{x_2},...,\mathbf{x_{t-1}},\mathbf{x_{t}})$ And the out put is a single scalar $\hat{y}$ which is supposed to be the prediction of the target $y$.
	Let $G_{\theta}(\mathbf{x_i},\mathbf{h_i})$ the $\theta$ parameterized LSTM, we have:
	\begin{equation}
	\mathbf{h_{i+1}}=G_{\theta}(\mathbf{x_i},\mathbf{h_i}),
	\end{equation}iteratively. and when we have the $\mathbf{h_{t}}$, we calculate our final prediction $\hat{y}$ by a fully connected layer with parameter $\mathbf{W}$:
	\begin{equation}
	\hat{y}=\mathbf{W}\cdot\mathbf{h_t},
	\end{equation}
	and the loss function is:
	\begin{equation}
	\mathcal{L}=\frac{1}{n}\sum_{i=1}^{n}(y_i-\hat{y_i})^2.
	\end{equation}
	The training objective is to minimize $\mathcal{L}$.

@@

	\subsubsection{Vector Input}
	In order to better leverage the data to get better performance with our LSTM model, we take the same featues for LSTM as random forest. More specifically, $F_A$ for target $R_r(t)$ and $F_C$ for target $avgRTT_P(t)$.The results are are shown in \textbf{Table~\ref{tab:lstm-vector}}.

	In order to better leverage the data to get better performance with our LSTM model, we take the same featues for LSTM model as those for random forest model. More specifically, Features for target $R_r(t)$ and Features for target $avgRTT_P(t)$.The results are shown in \textbf{Fig.\ref{fig:lstm-task1-vector}}, \textbf{Fig.\ref{fig:lstm-task3-vector}} and \textbf{Tab.\ref{tab:lstm-vector}}.
	\begin{table}[tbp]
		\centering
		\begin{tabular}{lcc}
			\toprule
			{Target} &{$R_r(t)$}&{$avgRTT_p(t)$} \\
			\midrule
			\textit{data-1}& 0.5374 & 0.7310 \\
			\textit{data-2}& 0.4956& 0.7039     \\
			\textit{data-3-ws10}& 0.7649 & 0.7521 \\
			\textit{data-3-ws20}& 0.6513& 0.6901    \\
			\textit{data-3-ws50}& 0.5647& 0.7074  \\
			\bottomrule	
		\end{tabular}
		\caption{ Experiment results ($ABC$) of LSTM models with vector input. }
		\label{tab:lstm-vector}
	\end{table}
	\begin{figure*} 	
		\includegraphics[width=0.3\textwidth]{fig4/data-1-task-1-7030-lstm.pdf}
		\includegraphics[width=0.3\textwidth]{fig4/data-2-task-1-7030-lstm.pdf}
	\end{figure*} 
	\begin{figure*} 
		\includegraphics[width=0.3\textwidth]{fig4/data-3-ws10-task-1-7030-lstm.pdf}
		\includegraphics[width=0.3\textwidth]{fig4/data-3-ws20-task-1-7030-lstm.pdf}
		\includegraphics[width=0.3\textwidth]{fig4/data-3-ws50-task-1-7030-lstm.pdf}
		%\vspace{-20pt}  
		\caption{Relative prediction errors for $R_r(t)$ of LSTM models with vector input.}
		\label{fig:lstm-task1-vector}
		%\vspace{-10pt}  
	\end{figure*}
	\begin{figure*} 	
	\includegraphics[width=0.3\textwidth]{fig4/data-1-task-3-7030-lstm.pdf}
	\includegraphics[width=0.3\textwidth]{fig4/data-2-task-3-7030-lstm.pdf}
	\end{figure*} 
	\begin{figure*} 	
	\includegraphics[width=0.3\textwidth]{fig4/data-3-ws10-task-3-7030-lstm.pdf}
	\includegraphics[width=0.3\textwidth]{fig4/data-3-ws20-task-3-7030-lstm.pdf}
	\includegraphics[width=0.3\textwidth]{fig4/data-3-ws50-task-3-7030-lstm.pdf}
	%\vspace{-20pt}  
	\caption{Relative prediction errors for  $avgRTT_P(t)$ of LSTM models with vector input.}
	\label{fig:lstm-task3-vector}
	%\vspace{-10pt}  
	\end{figure*}

@@

	However, when it comes to target $avgRTT_p(t)$, although LSTM models do have better performance when fed with vector input rather than scalar input, their $ABC$s are still smaller than those of random forest. We believe it is because there is less temporal locality with reference to $avgRTT_p(t)$. The more data we observed before, the better we can predict $avgRTT_p(t)$ in the future. Thus LSTM model cannot maximize its ability in predicting temporal patterns.
	\section{Conclusions}
	In this paper, We study the problem of predicting internet path dynamics and performance with \texttt{traceroute} measurement and machine learning models. We reprocess one dataset in several more rational ways and conduct extensive experiments on the reprocessed dataset with random forest and LSTM model. We find that some features e.g., $R_r(t)$, is more likely to have temporal locality than other features like $avgRTT_p(t)$, and this insight can guide us to select the most suitable model facing such tasks in computer network scenarios.

	%\begin{acks}
	%	TODO
	%\end{acks}

	\bibliographystyle{ACM-Reference-Format}
	\bibliography{test-bib}

>>>>

	\begin{abstract}
		We study the problem of predicting Internet path dynamics and performance with \texttt{traceroute} measurement and machine learning models. In this work, we first point out the drawbacks of the data processing in one previous work \cite{wassermann2017predicting}, and reprocess the data in several more rational ways. Then data analysis and experiments are conducted on the reprocessed data with random forest model before evident temporal locality is discovered in the behavior of the paths. Inspired by this property, we further leverage Long Short-Term Memory (LSTM) model, one kind of recurrent neural networks (RNN), and get promising results. We make our reprocessed data and experiment code publicly available\footnote{{https://github.com/felixwzh/CN-group-project}}.
	\end{abstract}

@@

	\section{Introduction}
	Internet paths change frequently due to inter/intra-domain routing changes, load balancing, and even misconfigurations and failures\cite{paxson1996end}.
	%  V. Paxson. 1996. End-to-end Routing Behavior in the Internet. In Proceedings of the ACM SIGCOMM 1996 Conference (SIGCOMM ’96). ACM, New York, NY, USA.
	Some of these changes can seriously disrupt performance, causing longer round-trip times, congestion, or even loss of connectivity\cite{javed2013poiroot}.
	%U. Javed, I. Cunha, D. Cho"nes, E. Katz-Bassett, T. Anderson, and A. Krishnamurthy. 2013. PoiRoot: Investigating the Root Cause of Interdomain Path Changes. In Proceedings of the ACM SIGCOMM 2013 Conference on SIGCOMM (SIGCOMM ’13). ACM, New York, NY, USA, 183–194.
	Thus, it is of great significant to predict Internet path dynamics and performance.

	In the original paper\cite{wassermann2017predicting}, the authors focus on the problem of predicting Internet path changes and path performance using \texttt{traceroute} measurements. They use the recent route information of paths (route age for paths, route changes in the past, average RTT, etc) to do some predictions of paths. Their are three predict targets: $(i)$ the remaining life time of a route (i.e., the time before a route changes), $(ii)$ the number of route changes in a future timeslot, and $(iii)$ the average RTT  that the path $P$ will experience in the next \texttt{traceroute} measurement.

@@

	The dataset the authors provide with is a full week of Paris- \texttt{traceroute} measurements performed through the M-Lab open
	Internet measurement initiative\footnote{https://www.measurementlab.net/}. They observe more than 450,000 different paths sampled through Paris-\texttt{traceroute} measurements from more than 180 geo-distributed servers. However, most of the paths are not periodically sampled during this week. And only 2,346 paths have at least 100 \texttt{traceroute} measurements during the analyzed week. So the dataset contains corresponding information of these 2,346 paths.

	We find some drawbacks in their data process.(see section \ref{sec:dataAna}) And in our own work, firstly we reprocess the provided data in more reasonable ways and use the same method (random forest model) mentioned by the original paper\cite{wassermann2017predicting} to compare the results. Secondly, another machine learning method, LSTM model, is applied to get better performance.

	\section{Related Works}
	In the original paper\cite{wassermann2017predicting}, the problem they solve is to use random forest model to predict three targets for one path: $(i)$ the remaining life time of a route for one path at time $t$ (i.e., the time before a route changes), namely $R_r(t)$, $(ii)$ the number of route changes for one path $P$ in a future timeslot $T$, namely $rcP_T(t)$  and $(iii)$ the average RTT that the path $P$ will experience in the next \texttt{traceroute} measurement, namely $avgRTT_P(t)$.

@@

	And the last group of 44 features ,referred to as $F_C$ describes statistical properties of path latency, which is relavent to the prediction of the next \texttt{traceroute} measurement. They calculate the statistical properties of four RTT metrics (average, minimum, maximum and standard deviation) reported from each \texttt{traceroute} measurement.

@@

	\section{Data Analysis}
	\label{sec:dataAna}
	With careful review of the data used in the original paper, here we: $(i)$ show the authors had used the data incorrectly in the context of making predictions, and $(ii)$ reveal the temporal properties that the original method failed to employ but perfectly fits LSTM model (long short term memory neural networks), which we will later use to improve the performance.
	\subsection{Justification of Data Processing}
	The original paper \cite{wassermann2017predicting} made a major mistake on data employment: they implicitly used data from the future to predict events in the future, which is not possible. More specifically, all of their tests on the performance of \textit{NetPerfTrace} depend on statistics calculated out of data of all time, including the data they were to predict. Features like percentiles of route life, for example, take the route life of all routes during the whole week into account, including route $r_i$. However, these features still take part in predicting the residual life of $r_i$. It resembles guessing something you already knew.

@@

	No more solid evidence could be provided to prove the periodicity of these patterns because pattern itself is hard to define. However, these results are already enough to make us hopeful about the performance of LSTM, which are expert in capturing temporal dependencies, no-matter long-term ones or short-term ones.

@@

	\subsection{Long Short-Term Memory}
	In order to leverage the prediction targets' character of temporal locality, we further use Long Short-Term Memory (LSTM) to do time series modeling. More formally, we take $t-1$ most recent \texttt{traceroute} measurements features $(\mathbf{x_1},\mathbf{x_2},...,\mathbf{x_{t-2}},\mathbf{x_{t-1}})$and the current \texttt{traceroute} measurement feature $\mathbf{x_t}$ together as the input, i.e., $\mathbf{X}=(\mathbf{x_1},\mathbf{x_2},...,\mathbf{x_{t-1}},\mathbf{x_{t}})$. And the output is a single scalar $\hat{y}$ which is supposed to be the prediction of the target $y$.
	Let $G_{\theta}(\mathbf{x_i},\mathbf{h_i})$ the $\theta$ parameterized LSTM, we have:
	\begin{equation}
	\mathbf{h_{i+1}}=G_{\theta}(\mathbf{x_i},\mathbf{h_i}),
	\end{equation}iteratively. and when we have the $\mathbf{h_{t}}$, we calculate our final prediction $\hat{y}$ by a fully connected layer with parameter $\mathbf{W}$:
	\begin{equation}
	\hat{y}=\mathbf{W}\cdot\mathbf{h_t},
	\end{equation}
	and the loss function is:
	\begin{equation}
	\mathcal{L}=\frac{1}{n}\sum_{i=1}^{n}(y_i-\hat{y_i})^2.
	\end{equation}
	The training objective is to minimize $\mathcal{L}$.

@@

	\subsubsection{Vector Input}
	%In order to better leverage the data to get better performance with our LSTM model, we take the same features for LSTM as random forest. More specifically, $F_A$ for target $R_r(t)$ and $F_C$ for target $avgRTT_P(t)$.The results are are shown in \textbf{Table~\ref{tab:lstm-vector}}.

	In order to better leverage the data to get better performance with our LSTM model, we take the same featues for LSTM model as those for random forest model. More specifically, Features for target $R_r(t)$ and features for target $avgRTT_P(t)$. The results are shown in \textbf{Fig.\ref{fig:lstm-task1-vector}}, \textbf{Fig.\ref{fig:lstm-task3-vector}} and \textbf{Tab.\ref{tab:lstm-vector}}.
	\begin{table}[tbp]
		\centering
		\begin{tabular}{lcc}
			\toprule
			{Target} &{$R_r(t)$}&{$avgRTT_p(t)$} \\
			\midrule
			\textit{data-1}& 0.5374 & 0.7310 \\
			\textit{data-2}& 0.4956& 0.7039     \\
			\textit{data-3-ws10}& 0.7649 & 0.7521 \\
			\textit{data-3-ws20}& 0.6513& 0.6901    \\
			\textit{data-3-ws50}& 0.5647& 0.7074  \\
			\bottomrule	
		\end{tabular}
		\caption{ Experiment results ($ABC$) of LSTM models with vector input. }
		\label{tab:lstm-vector}
	\end{table}
	\begin{figure*} 	
		\includegraphics[width=0.3\textwidth]{fig4/data-1-task-1-7030-lstm.pdf}
		\includegraphics[width=0.3\textwidth]{fig4/data-2-task-1-7030-lstm.pdf}
	\end{figure*} 
	\begin{figure*} 
		\includegraphics[width=0.3\textwidth]{fig4/data-3-ws10-task-1-7030-lstm.pdf}
		\includegraphics[width=0.3\textwidth]{fig4/data-3-ws20-task-1-7030-lstm.pdf}
		\includegraphics[width=0.3\textwidth]{fig4/data-3-ws50-task-1-7030-lstm.pdf}
		%\vspace{-20pt}  
		\caption{Relative prediction errors for $R_r(t)$ of LSTM models with vector input.}
		\label{fig:lstm-task1-vector}
		%\vspace{-10pt}  
	\end{figure*}
	\begin{figure*} 	
	\includegraphics[width=0.3\textwidth]{fig4/data-1-task-3-7030-lstm.pdf}
	\includegraphics[width=0.3\textwidth]{fig4/data-2-task-3-7030-lstm.pdf}
	\end{figure*} 
	\begin{figure*} 	
	\includegraphics[width=0.3\textwidth]{fig4/data-3-ws10-task-3-7030-lstm.pdf}
	\includegraphics[width=0.3\textwidth]{fig4/data-3-ws20-task-3-7030-lstm.pdf}
	\includegraphics[width=0.3\textwidth]{fig4/data-3-ws50-task-3-7030-lstm.pdf}
	%\vspace{-20pt}  
	\caption{Relative prediction errors for  $avgRTT_P(t)$ of LSTM models with vector input.}
	\label{fig:lstm-task3-vector}
	%\vspace{-10pt}  
	\end{figure*}

@@

	However, when it comes to target $avgRTT_p(t)$, although LSTM models do have better performance when fed with vector input rather than scalar input, their $ABC$s are still smaller than those of random forest. We believe it is because there is less temporal locality with reference to $avgRTT_p(t)$. The more data we observed before, the better we can predict $avgRTT_p(t)$ in the future. Thus LSTM model cannot maximize its ability in predicting temporal patterns.
	\section{Conclusions}
	In this paper, we study the problem of predicting Internet path dynamics and performance with \texttt{traceroute} measurement and machine learning models. We reprocess one dataset in several more rational ways and conduct extensive experiments on the reprocessed dataset with random forest and LSTM model. It is found that some features, e.g., $R_r(t)$, are more likely to have temporal locality than other features like $avgRTT_p(t)$, and this insight can guide us to select the most suitable model facing such tasks in computer network scenarios.

	\begin{acks}
	This work is successfully produced thanks to the kind guidance from and discussion with Prof. Linghe Kong. His revealed us common shortcomings of ML-CN works and inspired us to build models that suits \emph{network} better.
	\end{acks}

	{\small
		\bibliographystyle{ACM-Reference-Format}
		\bibliography{bibliography} 	
	}
